Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-08-09 surfaces-write-up/surface-write-up.tex
added response to todo
removed some todos

\todo{Make sure to include $u^d$ everywhere necessary}
\todo{Make sure to index starting from 0 in $\bp^m$ section}

@@

\section{Introduction}
\label{sec:intro}
For any Weil $\bq$-divisor $D$ on a rational surface $X$, the graded
\textbf{canonical ring} is $R(X, D) := \bigoplus_{d \geq 0} u^d H^0(X, \lfloor dD \rfloor)$, where $u$ is a dummy variable to keep
track of degree.
We study the generators and relations of canonical
rings for projective spaces $\bp^m$ and Hirzebruch surfaces $\hirz_m$ \todo{Peter: Bad to have $m$ as both the dimension of $\bp$ and the coefficient for the hirzebruch surface?}.
When $D$ is a general $\bq$-divisor on $\bp^m$ or
$\hirz_m$, we give bounds on the generators and relations of
$R(X, D)$. In particular, we give a complete description \todo{Peter: The phrase ``complete description'' is ambiguous and might be objectionable to a reader.  Is ``minimal presentation'' better?} of the
canonical ring when $X = \bp^m$ and $D$ is any effective
$\bq$-divisor.

@@

Another closely related topic is the Hassett-Keel program \cite{hassett:classical-and-minimal-models}, which aims to describe log canonical models of the form \todo{Robin: define $\mathscr{M}_g$? Not sure what this is.  \\
Peter: Isn't it defined in the line below?  (Maybe it was added after your comment)}
\begin{align*}
	\overline {\mathscr M}_g(\alpha) := \bigoplus_{d \geq 0} u^d H^0 \left(
	\overline {\mathscr M}_g, \lfloor d K_{\overline{\mathscr M}_g} +
	\alpha\delta \rfloor  \right) 
\end{align*}
in terms of certain moduli spaces, where $\overline {\mathscr M_g}$ is the
moduli space of stable genus $g$ curves.
Moreover, $\bq$-divisors on surfaces also naturally
appear when considering the canonical ring of surfaces of Kodaira dimension 1.
A surface has Kodaira dimension 1 if and only if it is an elliptic surface \cite[p. 244]{
barthHPV:compactComplexSurfaces}, and
the canonical rings of elliptic surfaces most naturally described in the setting of
$\bq$-divisors \cite[Chapter V, Theorem 12.1]{barthHPV:compactComplexSurfaces}.

@@

Write
\begin{align*}
	D = \sum_{i=0}^{n}\alpha_i D_i.
\end{align*}
where $n \in \bz$, $\alpha_i \in \bq$, and $\deg D_i = \bida_i$. Let $D_i$ be
Cartier divisors such that $D_i = V(f_i)$.
We shall further assume that (possibly after reordering) $f_0,\ldots, f_{m}$ form a
basis for degree one rational functions on $\bp^m$.  \todo{Peter: This assumption is inconsistent with a comment made later (see next todo); later, we allow for there to be only one $D_i$ ($D= D_i$), which contradicts $m>0$.}

@@

Next, we define an ordering on these vectors, which will be used to give a presentation for $R_D$. Although it is not
stated explicitly later, in the case that $D = \alpha_1 D_1$ \todo{Peter: Inconsistent with previous assumption.  To be clear it doesn't matter for any proofs, because that assumption is not necessary until after this section.  I suggest we find some way to move that assumption from earlier until after the effective case.} is supported on a single hypersurface, the relations in Theorem ~\ref{thm:proj-one-point} form a reduced grobner basis with respect to this ordering.

@@

Aaron: I agree that argument is obvious. I was more referring to the analog
of proposition ~\ref{prop:pm-span-and-basis}, that is, how to show
every numerator is a polynomial in $u,v,z,w$. This would probably be
an argument that takes a couple paragraphs.
}

@@

\noindent
lying in some degree $d \in \mathbb{N}$
with $c_j \ge -\alpha_j d$ (for all $j$) satisfying $\bida_i + \sum_{j = 1}
^n \bida_j c_j = 0$ and $\bidb_i + \sum_{j=1}^n \bidb_j c_j = 0$.
\todo{Aaron: The next two paragraphs are missing a lot of $u^d$ degree prefixes. Can one of you add them in and go through the proof
to make sure there are no others missing?\\
Peter: I added a bunch.  Should be better (although quite possibly missed a few more).}
Specifically, $\beta_i$,
is the polynomial $\beta_i(z_1,z_2,z_3,z_4)$ so that
$\beta_i(z_1,z_2,z_3,z_4) - z_i \in \ker \psi$. Such an element
$\beta$ exists and is unique because rational functions
whose numerators are polynomials in $f_1,f_2,f_3,f_4$ form a basis
of all rational functions in $R_D$. Furthermore, these
generate all relations, since they allow us to reduce any monomial $u^d\prod_{j =
1}^n z_j^{r_j}$ to the canonical form where $r_j = - \lfloor d \alpha_i
\rfloor$
whenever $j > 4$.  To bound the degree of generation of these relations, we bound the degree of generation of the ideal $(\beta_i  -z_i) \cap \ker(\psi)$ for each $i$.

>>>>

@@

\section{Introduction}
\label{sec:intro}
For any Weil $\bq$-divisor $D$ on a rational surface $X$, the graded
\textbf{canonical ring} is $R(X, D) := \bigoplus_{d \geq 0} u^d H^0(X, \lfloor dD \rfloor)$, where $u$ is a dummy variable to keep
track of degree.
We study the generators and relations of canonical
rings for projective spaces $\bp^m$ and Hirzebruch surfaces $\hirz_m$ \todo{Peter: Bad to have $m$ as both the dimension of $\bp$ and the coefficient for the hirzebruch surface?

Aaron: It doesn't seem like this would cause any confusing to me. They're always discussed separately.
}.
When $D$ is a general $\bq$-divisor on $\bp^m$ or
$\hirz_m$, we give bounds on the generators and relations of
$R(X, D)$. In particular, we give a minimal presentation of the
canonical ring when $X = \bp^m$ and $D$ is any effective
$\bq$-divisor.

@@

Another closely related topic is the Hassett-Keel program \cite{hassett:classical-and-minimal-models}, which aims to describe log canonical models of the form 
\begin{align*}
	\overline {\mathscr M}_g(\alpha) := \bigoplus_{d \geq 0} u^d H^0 \left(
	\overline {\mathscr M}_g, \lfloor d K_{\overline{\mathscr M}_g} +
	\alpha\delta \rfloor  \right) 
\end{align*}
in terms of certain moduli spaces, where $\overline {\mathscr M_g}$ is the
moduli space of stable genus $g$ curves.
Moreover, $\bq$-divisors on surfaces also naturally
appear when considering the canonical ring of surfaces of Kodaira dimension 1.
A surface has Kodaira dimension 1 if and only if it is an elliptic surface \cite[p. 244]{
barthHPV:compactComplexSurfaces}, and
the canonical rings of elliptic surfaces most naturally described in the setting of
$\bq$-divisors \cite[Chapter V, Theorem 12.1]{barthHPV:compactComplexSurfaces}.

@@

Write
\begin{align*}
	D = \sum_{i=0}^{n}\alpha_i D_i.
\end{align*}
where $n \in \bz$, $\alpha_i \in \bq$, and $\deg D_i = \bida_i$. Let $D_i$ be
Cartier divisors such that $D_i = V(f_i)$.
We shall further assume for this subsection that (possibly after reordering) $f_0,\ldots, f_{m}$ form a
basis for degree one rational functions on $\bp^m$.

@@

Next, we define an ordering on these vectors, which will be used to give a presentation for $R_D$. Although it is not
stated explicitly later, in the case that $D = \alpha_1 D_1$ is supported on a single hypersurface, the relations in
Theorem ~\ref{thm:proj-one-point} form a reduced grobner basis with respect to
this ordering.

@@

Aaron: I agree that argument is obvious. I was more referring to the analog
of proposition ~\ref{prop:pm-span-and-basis}, that is, how to show
every numerator is a polynomial in $u,v,z,w$. This would probably be
an argument that takes a couple paragraphs.
\\
Peter: I see what you mean.  What was the argument for that again.  It was some exact sequence thing you came up with right?
}

@@

\noindent
lying in some degree $d \in \mathbb{N}$
with $c_j \ge -\alpha_j d$ (for all $j$) satisfying $\bida_i + \sum_{j = 1}
^n \bida_j c_j = 0$ and $\bidb_i + \sum_{j=1}^n \bidb_j c_j = 0$.
Specifically, $\beta_i$,
is the polynomial $\beta_i(z_1,z_2,z_3,z_4)$ so that
$\beta_i(z_1,z_2,z_3,z_4) - z_i \in \ker \psi$. Such an element
$\beta$ exists and is unique because rational functions
whose numerators are polynomials in $f_1,f_2,f_3,f_4$ form a basis
of all rational functions in $R_D$. Furthermore, these
generate all relations, since they allow us to reduce any monomial $u^d\prod_{j =
1}^n z_j^{r_j}$ to the canonical form where $r_j = - \lfloor d \alpha_i
\rfloor$
whenever $j > 4$.  To bound the degree of generation of these relations, we bound the degree of generation of the ideal $(\beta_i  -z_i) \cap \ker(\psi)$ for each $i$.
